---
title: "2019 KHPC Analytics"
author: "Cornelius Tanui"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
    vertical_layout: scroll
    # source_code: embed
    logo: logo.jpg
    theme: lumen
date: "`r Sys.Date()`"
runtime: shiny
---

```{=html}
<!--
# Customizing HTML Appearance
-->
```
```{=html}
<style>
.navbar {
background-color:#438D80;
border-color: #15317E;
}

.navbar-brand {
color: #000000! important;
}

.section.sidebar {
background-color:#438D80;
border-color: #15317E;
color: #000000! important;
}

</style>
```
```{r Clear Evrironment, include = FALSE}
gc()
graphics.off()
rm(list = ls(all.names = TRUE))
```

```{r Load Packages, include = FALSE}
# load Packages
library(kableExtra)
library(MASS)
library(Matrix)
library(lattice)
library(mgcv)
library(nlme)
library(tidyverse)
library(flexdashboard)
library(shiny)
library(plotly)
library(DT)
library(ggplot2)
library(ggpmisc)
library(gmodels)
library(magrittr)
# devtools::install_github("Shelmith-Kariuki/rKenyaCensus")
# library(rKenyaCensus)
```

# Home

## Row {.tabset .tabset-fade}

### Introduction

This secondary data from Kenya Household and Population Census ([KPHC](https://www.knbs.or.ke/)) of 2019, made available in R by [Shelmith Kariuki](https://shelkariuki.netlify.app/#about) through her elegantly written [`rKenyaCensus`](https://github.com/Shelmith-Kariuki/rKenyaCensus) package.

This Shiny application uses [`flexdashboard`](https://rstudio.github.io/flexdashboard/index.html) framework, and [plotly](https://plotly.com/) graphing library's [`ggplotly`](https://www.rdocumentation.org/packages/plotly/versions/4.10.0/topics/ggplotly) function for interactivity of [ggplot2](https://ggplot2.tidyverse.org/) objects.

### Data

Data files available for public use;

```{r Describe data}
data_description<- rKenyaCensus::DataCatalogue %>% 
  as.data.frame()

data_description %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "400px")

```

# Dashboards

## Row {.tabset .tabset-fade}

### Demographics

```{r Socio-demographics}
# Distribution of Population by Age, Sex, County and Sub- County

# Remove unnecessary records 
pp_age_sex_county <- rKenyaCensus::V3_T2.3 %>% 
  filter(SubCounty != "ALL" & 
           Age != "Total" & 
           Age != "Not Stated" &
           !grepl("-", Age)
         )

# For age above 100 years, replace with 100 
pp_age_sex_county[pp_age_sex_county$Age == "100+", ]$Age <- "100"

pp_age_sex_county$Age <- as.numeric(pp_age_sex_county$Age)

# Transform data to long
male <- pp_age_sex_county %>% 
  select(County, SubCounty, Age, Male) %>% 
  rename(Population = Male) %>% 
  mutate(Gender = "Male")

female <- pp_age_sex_county %>% 
  select(County, SubCounty, Age, Female)%>% 
  rename(Population = Female)%>% 
  mutate(Gender = "Female")

total <- pp_age_sex_county %>% 
  select(County, SubCounty, Age, Total)%>% 
  rename(Population = Total)%>% 
  mutate(Gender = "Total")

pp_age_sex_county_long <- bind_rows(male, female, total)

# plot data (population  by age and sex)
plot1 <- pp_age_sex_county_long %>% 
  ggplot(aes(x = Age, y = Population)) +
  geom_point() +
  geom_jitter(color = "#009999") +
  labs(x = "Age",
       y = "Population",
       title = "Population by Age and Sex",
       caption = "Source: KHPC2019") +
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"))+
  facet_wrap(. ~ Gender)

# Animate plot 
plot1 <- ggplotly(plot1)

# plot1
```

### Health

```{r Health}

```
